### 2023-03-16 chapters/paper_review/04b_analysis.tex
Change reference keys for review paper (WIP)

In \cite{pasad2021} wav2vec~2.0 representations were analyzed with
respect to their acoustic-linguistic information content at different
network layers. Three different mechanisms were used for this
purpose. The first of these is canonical correlation analysis (CCA),
which computes similarity scores between two continuous vectors
based on the maximum correlation of their linear projections.  These
can be used to judge the similarity of embeddings at different layers
with each other, with standard acoustic representations such as mel
filterbank features, or word embeddings derived from text. The second
method clusters continuous representation vectors and computes the
discrete mutual information between cluster IDs and phone or word
labels. The third method involves probing tasks: representation
vectors extracted from the network are used to perform simple
downstream tasks, in particular determining whether two acoustic segments
correspond to the same word, and a standard benchmark of 11 word
similarity tasks~\cite{faruqi14}. These are mostly used to gauge the
amount of lexical information present in the embeddings.  Using this
battery of tests the authors compared pre-trained models of varying
sizes as well as models fine-tuned for ASR. They found that pre-trained
models show an autoencoder-style behavior, with early layers showing
strong similarity with input features, intermediate layers diverging
more, and final layers reverting back to higher similarity with input
features and early layers. Generally, the earlier layers in wav2vec~2.0 
models encode acoustic information. The next set of layers encodes
phonetic class information, followed by word meaning information,
before reverting back to encoding phonetic/acoustic information. Thus,
extracting representations from the last layers for tasks that require
phonetic or word-related information may not be the best
strategy. Indeed, the authors of \cite{baevski2021unsupervised} show that a
phone classifier trained on each of the 24 frozen layers of a wav2vec~2.0 model
showed the lowest phone error rates
for layers 10--21 and higher error rates for the other layers. 
\cite{pasad2021} further show that fine-tuning the pre-trained model with a
character-level CTC training criterion changes the behavior of the
last layers (especially the final two layers), breaking the
autoencoder-style behavior and focusing the information encoded in the
last layers on orthographic-phonetic and word information.

@@

Self-supervised representations may encode other information besides phonetic
classes or words, for example, channel, language, speaker, and sentiment
information. 
\edit{It is shown that the per-utterance mean of CPC features captures speaker information to a large extent\cite{van2021analyzing}.}
Location of information pertaining to speakers vs.\ language classes was
analyzed in \cite{ling20odyssey} for a 12-layer BERTphone model. This model
combines a self-supervised masked reconstruction loss with a phone-based CTC
loss to produce representations	for speaker recognition and language
identification. By	analyzing the weights	of a linear combination	of layer
representations for these two downstream tasks, it was	shown that language
recognition draws on representation	from higher layers (peaking at layer~10)
whereas speaker recognition benefited from layers at positions 6, 9, and 12.
This may indicate that language recognition relies more on higher-level
phonetic information whereas speaker recognition uses a combination of
acoustic and phonetic information. In a recent study~\cite{chen2021wavlm} the
same technique was used to identify layer contributions for the downstream
SUPERB benchmark tasks in the WavLM model. For a smaller model (95M parameters)
it was again confirmed that lower layers encode speaker-related information
necessary for speaker diarization and verification whereas higher layers encode
phonetic and semantic information. Another study~\cite{wang2021layersup} used
explicit self-supervised loss at the intermediate layers rather than just the
output layer of a HuBERT model in order to enforce better learning of phonetic
information. The resulting model was indeed better at downstream tasks
requiring information about phonetic content, such as phone recognition, ASR,
and keyword spotting, but worse at speaker-related tasks like speaker
diarization and verification.

@@

How does the performance of self-supervised models change in relation to the
amount of training data, and in relation to the size (number of parameters) of
the model?
Several studies have demonstrated better downstream performance when using
larger datasets~\cite{rivi20, kawakami2020learning, chen2021wavlm}. For
example,
\cite{kawakami2020learning} compared representations learned by a bidirectional
CPC model from the standard 960 hour LS corpus and a corpus of 8,000 hours of
diverse speech from multiple sources.\textsuperscript{\ref{footnote:CPC-8k}}
Not surprisingly, an ASR model trained on top of these representations
performed better when representations were learned from the larger dataset. 
Although the precise relationship between data size and performance has not
been quantified, we can assume that it follows a law of diminishing returns (or
power law), 
similar to observations for most data-intensive machine learning tasks.
In addition to the size of the dataset, the diversity of the data also seems
to play a role, although this was not quantified in this study. However, recent
experiments with larger and more diverse data collections~\cite{chen2021wavlm}
confirm this assumption, as do 
explicit investigations of domain shift robustness (see
\cref{subsec:robustness} below).

\edit{The relation between model sizes and downstream performances have also been investigated \cite{pue2021scaling,versteegh2015zero}.}
Using the Mockingjay 
model~\cite{liu2020mockingjay}, the authors in \cite{pue2021scaling} attempt to establish a relationship between
model size and self-supervised \ensuremath{L_1} loss and demonstrate that it approximately
follows a power law. Model size and accuracy on downstream phone
classification and speaker recognition tasks are positively correlated but do not
exactly follow a power law; rather, the accuracy saturates as models increase in
size, possibly due to the lack of a corresponding expansion in training data
size.

>>>>

In \cite{pasad_layerwise_2021} wav2vec~2.0 representations were analyzed with
respect to their acoustic-linguistic information content at different
network layers. Three different mechanisms were used for this
purpose. The first of these is canonical correlation analysis (CCA),
which computes similarity scores between two continuous vectors
based on the maximum correlation of their linear projections.  These
can be used to judge the similarity of embeddings at different layers
with each other, with standard acoustic representations such as mel
filterbank features, or word embeddings derived from text. The second
method clusters continuous representation vectors and computes the
discrete mutual information between cluster IDs and phone or word
labels. The third method involves probing tasks: representation
vectors extracted from the network are used to perform simple
downstream tasks, in particular determining whether two acoustic segments
correspond to the same word, and a standard benchmark of 11 word
similarity tasks~\cite{faruqi14}. These are mostly used to gauge the
amount of lexical information present in the embeddings.  Using this
battery of tests the authors compared pre-trained models of varying
sizes as well as models fine-tuned for ASR. They found that pre-trained
models show an autoencoder-style behavior, with early layers showing
strong similarity with input features, intermediate layers diverging
more, and final layers reverting back to higher similarity with input
features and early layers. Generally, the earlier layers in wav2vec~2.0 
models encode acoustic information. The next set of layers encodes
phonetic class information, followed by word meaning information,
before reverting back to encoding phonetic/acoustic information. Thus,
extracting representations from the last layers for tasks that require
phonetic or word-related information may not be the best
strategy. Indeed, the authors of \cite{baevski2021unsupervised} show that a
phone classifier trained on each of the 24 frozen layers of a wav2vec~2.0 model
showed the lowest phone error rates
for layers 10--21 and higher error rates for the other layers. 
\cite{pasad_layerwise_2021} further show that fine-tuning the pre-trained model with a
character-level CTC training criterion changes the behavior of the
last layers (especially the final two layers), breaking the
autoencoder-style behavior and focusing the information encoded in the
last layers on orthographic-phonetic and word information.

@@

Self-supervised representations may encode other information besides phonetic
classes or words, for example, channel, language, speaker, and sentiment
information. 
\edit{It is shown that the per-utterance mean of CPC features captures speaker information to a large extent\cite{van2021analyzing}.}
Location of information pertaining to speakers vs.\ language classes was
analyzed in \cite{ling20odyssey} for a 12-layer BERTphone model. This model
combines a self-supervised masked reconstruction loss with a phone-based CTC
loss to produce representations	for speaker recognition and language
identification. By	analyzing the weights	of a linear combination	of layer
representations for these two downstream tasks, it was	shown that language
recognition draws on representation	from higher layers (peaking at layer~10)
whereas speaker recognition benefited from layers at positions 6, 9, and 12.
This may indicate that language recognition relies more on higher-level
phonetic information whereas speaker recognition uses a combination of
acoustic and phonetic information. In a recent study~\cite{chen_wavlm_2021} the
same technique was used to identify layer contributions for the downstream
SUPERB benchmark tasks in the WavLM model. For a smaller model (95M parameters)
it was again confirmed that lower layers encode speaker-related information
necessary for speaker diarization and verification whereas higher layers encode
phonetic and semantic information. Another study~\cite{wang2021layersup} used
explicit self-supervised loss at the intermediate layers rather than just the
output layer of a HuBERT model in order to enforce better learning of phonetic
information. The resulting model was indeed better at downstream tasks
requiring information about phonetic content, such as phone recognition, ASR,
and keyword spotting, but worse at speaker-related tasks like speaker
diarization and verification.

@@

How does the performance of self-supervised models change in relation to the
amount of training data, and in relation to the size (number of parameters) of
the model?
Several studies have demonstrated better downstream performance when using
larger datasets~\cite{rivi20, kawakami2020learning, chen_wavlm_2021}. For
example,
\cite{kawakami2020learning} compared representations learned by a bidirectional
CPC model from the standard 960 hour LS corpus and a corpus of 8,000 hours of
diverse speech from multiple sources.\textsuperscript{\ref{footnote:CPC-8k}}
Not surprisingly, an ASR model trained on top of these representations
performed better when representations were learned from the larger dataset. 
Although the precise relationship between data size and performance has not
been quantified, we can assume that it follows a law of diminishing returns (or
power law), 
similar to observations for most data-intensive machine learning tasks.
In addition to the size of the dataset, the diversity of the data also seems
to play a role, although this was not quantified in this study. However, recent
experiments with larger and more diverse data collections~\cite{chen_wavlm_2021}
confirm this assumption, as do 
explicit investigations of domain shift robustness (see
\cref{subsec:robustness} below).

\edit{The relation between model sizes and downstream performances have also been investigated \cite{pue2021scaling,versteegh2015zero}.}
Using the Mockingjay 
model~\cite{liu_mockingjay_2020}, the authors in \cite{pue2021scaling} attempt to establish a relationship between
model size and self-supervised \ensuremath{L_1} loss and demonstrate that it approximately
follows a power law. Model size and accuracy on downstream phone
classification and speaker recognition tasks are positively correlated but do not
exactly follow a power law; rather, the accuracy saturates as models increase in
size, possibly due to the lack of a corresponding expansion in training data
size.

### 2023-03-16 chapters/paper_review/05_zero.tex
Change reference keys for review paper (WIP)

\begin{table*}[ht]
\centering
    {\renewcommand*\arraystretch{1.4}}
\caption{
Unsupervised ASR.
TIMIT numbers are phoneme error rates (PER), while the numbers for
LibriSpeech are word error rates (WER).
SWC $=$ spoken word classifier, ST $=$ speech translation.
All speech and text are in English if not specified. 
\textcolor{black}{
The references in the table are sorted according to the date of publication.}
}
\label{fig:unsupervisedASR}
\resizebox{0.99\textwidth}{!}{
\begin{tabular}{cllllll}
    \toprule
Reference & Speech representation & Speech segmentation  & Text token/representation & Mapping approach & Refinement  & Results  \\ 
    \midrule
\cite{liu2018completely}& Audio word2vec~\cite{Wang2018SegAudioWord2Vec} & Oracle & Phoneme & \edit{Adversarial Training~\cite{gulrajani2017improved}} & - & TIMIT (PER): 63.6\%  \\ %Submitted on 1 Apr 2018
  \midrule
  \cite{chung2018unsupervised} & Speech2vec~\cite{chung2018speech2vec} & BES-GMM~\cite{Kamper17BESGMM} &  Word2Vec & \edit{Adversarial Training~\cite{conneau2017word}} & Self-training & SWC (Acc): 10.9\%  \\ %Submitted on 18 May 2018
%https://arxiv.org/pdf/1805.07467.pdf %The text embeddings were obtained by training Word2Vec on the transcriptions using the fastText implementation without subword information [3].
  \midrule

@@

% how to represent audio
For unsupervised ASR to be feasible, the common idea is to make the speech and text units close to each other.
For the text side, word sequences can be transformed into phoneme sequences if a lexicon is available. 
On the other hand, we must first convert the speech signal into something close to phonemes. 
To achieve that, most studies on unsupervised ASR use a phoneme
segmentation module before the generator to segment utterances into
phoneme-level
segments~\cite{liu2018completely,chen2019completely,yeh2018unsupervised}. 
A representation vector or a token then represents each phoneme-level segment. 
It is easier for the generator to map each segment-level representation 
or token to the correct phoneme when the representation or token is highly correlated to the phonemes.
Wav2vec-U~\cite{baevski2021unsupervised} selects the input feature from different layers of wave2vec~2.0~\cite{baevski2020wav2vec}.
The selection criterion is based on analysis of the phonetic information in each layer.
\textcolor{black}{
If a universal phone recognizer trained from a diverse set of languages is available, it is another way to transcribe speech into phone-level tokens~\cite{Deciphering2022}.}
Another series of work is to transform a word into a word embedding. 
\cite{chung2018unsupervised,chung2019towards} use adversarial training to map the word-level
speech embedding space~\cite{chung2018speech2vec} to the word embedding space
and achieve promising performance on spoken word classification, speech
translation, and spoken word retrieval.
\Cref{fig:unsupervisedASR} summarizes the various ways to segment speech and represent speech and text in each reference.

%Hung-yi Lee: I comment this paragraph first.
\begin{comment}
%Hung-yi Lee: This paragraph is not finished yet.
\TNS{if we are limited on space,we can consider removing some of the detail of the various algorithms}
\TNS{split into 2 sentences, also have a connection to the previous para}
\cite{liu2018completely} is the first work to successfully achieve unsupervised
phoneme recognition by first clustering phoneme-level
embeddings~\cite{Chung2016AudioWord2Vec,Wang2018SegAudioWord2Vec} into a set of
tokens and then using GAN to learn the mapping relationship between tokens and
phonemes. 
However, in \cite{liu2018completely}, oracle phoneme segmentation is still
required to get reasonable performance. 
\cite{chen2019completely,yeh2018unsupervised} uses gate activation signals
(GAS)~\cite{wang2017gate} to get the phoneme-level segmentation, which is
unsupervised. 
%However, the loss term in~\cite{yeh2018unsupervised} includes an empirical average over all segments inside a logarithmic function, which can be biased if we sample this empirical average by a mini-batch average. \cite{liu2017unsupervised} therefore use an extremely large batch size during training to reduce the biasing problem.
%These problems are handled by another work~\cite{chen2019completely}.
%\cite{chen2019completely} removes the quantization step.
%Compared to~\cite{yeh2018unsupervised}, the discriminator in~\cite{chen2019completely} considers all possible information from the text dataset, not limited to the n-gram local statistics.
%Besides, \cite{chen2019completely} works well with a reasonable batch size, which makes it feasible when the computational resource is limited.
Wav2vec-U~\cite{baevski2021unsupervised} selects the input feature from
different layers of wave2vec 2.0~\cite{baevski2020wav2vec}, a self-supervised
model.
The selecting criterion is the PER by training a linear model in a supervised manner. 
%Should we mention this?
Wav2vec-U uses \kmeans{} to cluster the selected features, and the boundaries
are drawn whenever the clustered index changes.
%Wave2vec-U tends to generate more segments compared to the oracle phoneme-level segments. For example, their segmentation gives 0.93 precision, the ratio of the oracle phoneme-level boundaries being hit, and 0.38 recall, the ratio of the generated boundaries hit the oracle phoneme-level boundaries.
While Wav2vec-U merges the neighboring segments containing the same predicted
labels in each step of GAN training, this design can be viewed as refining the
segmentation implicitly.  
%Should we mention this?
%At generator output, the wav2vec-U will combine the neighboring predicted posterior containing the same predicted label into a predicted posterior.
\end{comment}

@@

An emerging technique uses a well-trained TTS system to generate
speech and text data from text-only data.
This technique is a sub-problem of the TTS-ASR approach formulated in
\eqref{eq:tts-asr} in which we fix the TTS system part and estimate only the ASR
parameters. 
For example, a huge amount of text resources can be obtained from the web
and document archives without corresponding audio data.
%Self-supervised learning in this paper mostly focuses on representation learning perspectives of such unpaired data.
The typical use case scenario of such a text resource for ASR is through \edit{the
language model}. 
We combine the ASR and language model via a noisy channel 
model~\cite{jelinek1997statistical}, a weighted finite state 
transducer~\cite{mohri2002weighted}, or shallow 
fusion~\cite{gulcehre2015using,chorowski2016towards}.
However, the progress of TTS systems boosted by deep 
learning~\cite{oord2016wavenet,shen2018natural} has inspired another interesting and
straightforward research direction: \edit{artificially creating paired text and audio
data $\{\hat{X}, Y\}$ with only text data~$Y$ by generating the corresponding
audio data~$\hat{X}$ with TTS.} %\begin{equation}
%    \{\hat{X}, Y\} \text{ where } \hat{X} = g_{\mathrm{tts}} (Y).
%\end{equation}
The most straightforward approach is to simply use multi-speaker TTS to
generate the waveform with various acoustic 
variations~\cite{li2018training,ueno2019multi,rosenberg2019speech,laptev2020you,huang2020using}.
The other approaches are based on the generation of high-level (more
linguistic) features instead of generating the waveform, e.g., encoder
features~\cite{hayashi2018back} and phoneme 
features~\cite{renduchintala18_interspeech,masumura20_interspeech}.
This approach is similar to the back-translation technique developed in \edit{neural machine translation}~\cite{sennrich2016improving}.
One benefit of the above data generation approaches is that it can be used to feed
unseen word or context phrases to end-to-end ASR.

@@

In this section, we describe the research directions of zero-resource speech
technologies by following the series of zero-resource speech challenges.
\begin{itemize}
	 \item \edit{Zero Resource Speech Challenge 2015~\cite{versteegh2015zero} mainly focused on building an acoustic model without using any
	 linguistic annotations based on subword unit modeling and spoken term
	 discovery tracks}.
	 \edit{For the subword unit modeling track, the ABX score for the within- and across-speaker
	 tasks was used as an evaluation metric.}
	 The spoken term discovery track used the normalized edit distance and coverage
	 scores in addition to the precision, recall, and F1 scores for types,
	 tokens, and boundaries.
    Both tracks were based on the English and Xitsonga languages.
	 \item The Zero Resource Speech Challenge 2017~\cite{dunbar2017zero} focused on 
	 unseen language and speaker aspects from the previous challenge. For
	 example, to demonstrate the robustness against unseen languages, the systems
	 were developed with English, French, and Mandarin and tested on
	 two ``surprise'' languages: German and Wolof.
	 Similarly, robustness against unseen speakers was demonstrated by varying
	 the amount of speech available for each speaker.
	 \item The Zero Resource Speech Challenge 2019~\cite{dunbar2019zero} extended a goal of
	 previous challenges by synthesizing speech without
	 text or phonetic labels but with acoustic units obtained using
	 zero-resource techniques.
	 The evaluation metrics were also extended to subjectively evaluate the
	 quality of synthesized speech, including its intelligibility, naturalness,
	 and speaker similarity.
	 \item The Zero Resource Speech Challenge 2020~\cite{dunbar2020zero} was based on two
	 tracks, revisiting previous challenges with different evaluation metrics. 
	 The first task revisited the 2019 challenge with low bit-rate subword
	 representations that optimize the quality of speech synthesis. The second
	 task revisited the 2017 challenge by focusing on the discovery of word-like
	 units from unsegmented raw speech.
	 \item The Zero Resource Speech Challenge 2021~\cite{nguyen2020zero}, the latest
	 \edit{challenge}, expanded the scope to include language modeling tasks.
	 In addition to phoneme-level ABX, the challenge includes lexical,
	 semantic, and syntactic evaluation metrics computed via a language model of
	 pseudo-acoustic labels.
\end{itemize}
These challenges have facilitated the tracking of technical trends in
zero-resource speech technologies.
For example, research directions thereof 
have expanded to various speech processing components to cover the entire
spoken dialogue systems.
To keep up with this expansion, the challenge has continued to develop
appropriate evaluation metrics for zero-resource scenarios.
Following the success of representation learning, baseline and challenge
techniques have shifted from purely generative 
models~\cite{ondel2016variational,heck2017feature}, deep 
autoencoders~\cite{tjandra19_interspeech,chorowski2019unsupervised}, and incorporation of
neural-network-based TTS/VC techniques~\cite{niekerk20b_interspeech} to
self-supervised learning~\cite{maekaku21_interspeech}.
The latest challenge included the visual modality, continuing the
expansion to include more aspects of human interaction.

@@

Textless NLP is a new research direction that leverages the progress mentioned
above in self-supervised speech representation learning to model language
directly from audio, bypassing the need for text or 
labels~\cite{GSLM,SpeechResynthesis_IS21,pGSLM,emotion_conversion}. 
Not only does this open the gate for language and dialect modeling without
orthographic rules, but it also offers the opportunity to model other
non-lexical information about how speech is delivered, e.g., speaker identity,
emotion, hesitation, interruptions. 
The generative spoken language model (GSLM)~\cite{GSLM} utilizes discrete
representations from wav2vec~2.0, HuBERT, and CPC algorithms as inputs to an
autoregressive language model trained by \edit{using the cross-entropy function} to maximize the
probability of predicting the next discrete speech token. A \edit{synthesis} module
follows the language model to produce speech waveforms given the generated
discrete speech units. The generated spoken continuations compete with
supervised generations and synthesis using a character language model in
subjective human evaluations. The model completes incomplete words
(pow[..] $\rightarrow$ POWER) and continues using words in the same general mood (dark $\rightarrow$
BLACKNESS)\footnote{https://speechbot.github.io/gslm/} and has been extended to
model and generate
dialogues~\cite{dialogue_GSLM}.\footnote{https://speechbot.github.io/dgslm/}
%\cite{}. 
Given its flexibility in modeling spoken content, the GSLM has been further extended
to jointly model content and prosody~\cite{pGSLM}. This prosodic-GSLM model
introduced a multistream causal Transformer, where the input and output layers
use multiple heads to model three channels:  discrete speech units, duration,
and quantized pitch. The prosodic-GSLM model jointly generates novel content
and prosody congruently in the expressive 
style of the prompt.\footnote{https://speechbot.github.io/pgslm/}
%\cite{}. 
Going one step further, \cite{emotion_conversion} used a speech emotion
conversion framework to modify the perceived emotion of a speech utterance
while preserving its lexical content and speaker identity. Other studies have
extended the idea of textless language processing or audio discrete representation to applications such as 
spoken question answering~\cite{spoken_qa_dual}, speech separation~\cite{shi2021discretization}, TTS~\cite{hayashi2020discretalk}, and speech-to-speech
translation~\cite{textless_translation}.

>>>>

\begin{table*}[ht]
\centering
    {\renewcommand*\arraystretch{1.4}}
\caption{
Unsupervised ASR.
TIMIT numbers are phoneme error rates (PER), while the numbers for
LibriSpeech are word error rates (WER).
SWC $=$ spoken word classifier, ST $=$ speech translation.
All speech and text are in English if not specified. 
\textcolor{black}{
The references in the table are sorted according to the date of publication.}
}
\label{fig:unsupervisedASR}
\resizebox{0.99\textwidth}{!}{
\begin{tabular}{cllllll}
    \toprule
Reference & Speech representation & Speech segmentation  & Text token/representation & Mapping approach & Refinement  & Results  \\ 
    \midrule
\cite{liu2018completely}& Audio word2vec~\cite{Wang2018SegAudioWord2Vec} & Oracle & Phoneme & \edit{Adversarial Training~\cite{gulrajani2017improved}} & - & TIMIT (PER): 63.6\%  \\ %Submitted on 1 Apr 2018
  \midrule
  \cite{chung2018unsupervised} & Speech2vec~\cite{chung_speech2vec_2018} & BES-GMM~\cite{kamper_segmental_2017} &  Word2Vec & \edit{Adversarial Training~\cite{conneau2017word}} & Self-training & SWC (Acc): 10.9\%  \\ %Submitted on 18 May 2018
%https://arxiv.org/pdf/1805.07467.pdf %The text embeddings were obtained by training Word2Vec on the transcriptions using the fastText implementation without subword information [3].
  \midrule

@@

% how to represent audio
For unsupervised ASR to be feasible, the common idea is to make the speech and text units close to each other.
For the text side, word sequences can be transformed into phoneme sequences if a lexicon is available. 
On the other hand, we must first convert the speech signal into something close to phonemes. 
To achieve that, most studies on unsupervised ASR use a phoneme
segmentation module before the generator to segment utterances into
phoneme-level
segments~\cite{liu2018completely,chen2019completely,yeh2018unsupervised}. 
A representation vector or a token then represents each phoneme-level segment. 
It is easier for the generator to map each segment-level representation 
or token to the correct phoneme when the representation or token is highly correlated to the phonemes.
Wav2vec-U~\cite{baevski2021unsupervised} selects the input feature from different layers of wave2vec~2.0~\cite{baevski2020wav2vec}.
The selection criterion is based on analysis of the phonetic information in each layer.
\textcolor{black}{
If a universal phone recognizer trained from a diverse set of languages is available, it is another way to transcribe speech into phone-level tokens~\cite{Deciphering2022}.}
Another series of work is to transform a word into a word embedding. 
\cite{chung2018unsupervised,chung2019towards} use adversarial training to map the word-level
speech embedding space~\cite{chung_speech2vec_2018} to the word embedding space
and achieve promising performance on spoken word classification, speech
translation, and spoken word retrieval.
\Cref{fig:unsupervisedASR} summarizes the various ways to segment speech and represent speech and text in each reference.

%Hung-yi Lee: I comment this paragraph first.
\begin{comment}
%Hung-yi Lee: This paragraph is not finished yet.
\TNS{if we are limited on space,we can consider removing some of the detail of the various algorithms}
\TNS{split into 2 sentences, also have a connection to the previous para}
\cite{liu2018completely} is the first work to successfully achieve unsupervised
phoneme recognition by first clustering phoneme-level
embeddings~\cite{chung_audio_2016,Wang2018SegAudioWord2Vec} into a set of
tokens and then using GAN to learn the mapping relationship between tokens and
phonemes. 
However, in \cite{liu2018completely}, oracle phoneme segmentation is still
required to get reasonable performance. 
\cite{chen2019completely,yeh2018unsupervised} uses gate activation signals
(GAS)~\cite{wang2017gate} to get the phoneme-level segmentation, which is
unsupervised. 
%However, the loss term in~\cite{yeh2018unsupervised} includes an empirical average over all segments inside a logarithmic function, which can be biased if we sample this empirical average by a mini-batch average. \cite{liu2017unsupervised} therefore use an extremely large batch size during training to reduce the biasing problem.
%These problems are handled by another work~\cite{chen2019completely}.
%\cite{chen2019completely} removes the quantization step.
%Compared to~\cite{yeh2018unsupervised}, the discriminator in~\cite{chen2019completely} considers all possible information from the text dataset, not limited to the n-gram local statistics.
%Besides, \cite{chen2019completely} works well with a reasonable batch size, which makes it feasible when the computational resource is limited.
Wav2vec-U~\cite{baevski2021unsupervised} selects the input feature from
different layers of wave2vec 2.0~\cite{baevski2020wav2vec}, a self-supervised
model.
The selecting criterion is the PER by training a linear model in a supervised manner. 
%Should we mention this?
Wav2vec-U uses \kmeans{} to cluster the selected features, and the boundaries
are drawn whenever the clustered index changes.
%Wave2vec-U tends to generate more segments compared to the oracle phoneme-level segments. For example, their segmentation gives 0.93 precision, the ratio of the oracle phoneme-level boundaries being hit, and 0.38 recall, the ratio of the generated boundaries hit the oracle phoneme-level boundaries.
While Wav2vec-U merges the neighboring segments containing the same predicted
labels in each step of GAN training, this design can be viewed as refining the
segmentation implicitly.  
%Should we mention this?
%At generator output, the wav2vec-U will combine the neighboring predicted posterior containing the same predicted label into a predicted posterior.
\end{comment}

@@

An emerging technique uses a well-trained TTS system to generate
speech and text data from text-only data.
This technique is a sub-problem of the TTS-ASR approach formulated in
\eqref{eq:tts-asr} in which we fix the TTS system part and estimate only the ASR
parameters. 
For example, a huge amount of text resources can be obtained from the web
and document archives without corresponding audio data.
%Self-supervised learning in this paper mostly focuses on representation learning perspectives of such unpaired data.
The typical use case scenario of such a text resource for ASR is through \edit{the
language model}. 
We combine the ASR and language model via a noisy channel 
model~\cite{jelinek1997statistical}, a weighted finite state 
transducer~\cite{mohri2002weighted}, or shallow 
fusion~\cite{gulcehre2015using,chorowski2016towards}.
However, the progress of TTS systems boosted by deep 
learning~\cite{oord_wavenet_2016,shen2018natural} has inspired another interesting and
straightforward research direction: \edit{artificially creating paired text and audio
data $\{\hat{X}, Y\}$ with only text data~$Y$ by generating the corresponding
audio data~$\hat{X}$ with TTS.} %\begin{equation}
%    \{\hat{X}, Y\} \text{ where } \hat{X} = g_{\mathrm{tts}} (Y).
%\end{equation}
The most straightforward approach is to simply use multi-speaker TTS to
generate the waveform with various acoustic 
variations~\cite{li2018training,ueno2019multi,rosenberg2019speech,laptev2020you,huang2020using}.
The other approaches are based on the generation of high-level (more
linguistic) features instead of generating the waveform, e.g., encoder
features~\cite{hayashi2018back} and phoneme 
features~\cite{renduchintala18_interspeech,masumura20_interspeech}.
This approach is similar to the back-translation technique developed in \edit{neural machine translation}~\cite{sennrich2016improving}.
One benefit of the above data generation approaches is that it can be used to feed
unseen word or context phrases to end-to-end ASR.

@@

In this section, we describe the research directions of zero-resource speech
technologies by following the series of zero-resource speech challenges.
\begin{itemize}
	 \item \edit{Zero Resource Speech Challenge 2015~\cite{versteegh2015zero} mainly focused on building an acoustic model without using any
	 linguistic annotations based on subword unit modeling and spoken term
	 discovery tracks}.
	 \edit{For the subword unit modeling track, the ABX score for the within- and across-speaker
	 tasks was used as an evaluation metric.}
	 The spoken term discovery track used the normalized edit distance and coverage
	 scores in addition to the precision, recall, and F1 scores for types,
	 tokens, and boundaries.
    Both tracks were based on the English and Xitsonga languages.
	 \item The Zero Resource Speech Challenge 2017~\cite{dunbar2017zero} focused on 
	 unseen language and speaker aspects from the previous challenge. For
	 example, to demonstrate the robustness against unseen languages, the systems
	 were developed with English, French, and Mandarin and tested on
	 two ``surprise'' languages: German and Wolof.
	 Similarly, robustness against unseen speakers was demonstrated by varying
	 the amount of speech available for each speaker.
	 \item The Zero Resource Speech Challenge 2019~\cite{dunbar2019zero} extended a goal of
	 previous challenges by synthesizing speech without
	 text or phonetic labels but with acoustic units obtained using
	 zero-resource techniques.
	 The evaluation metrics were also extended to subjectively evaluate the
	 quality of synthesized speech, including its intelligibility, naturalness,
	 and speaker similarity.
	 \item The Zero Resource Speech Challenge 2020~\cite{dunbar2020zero} was based on two
	 tracks, revisiting previous challenges with different evaluation metrics. 
	 The first task revisited the 2019 challenge with low bit-rate subword
	 representations that optimize the quality of speech synthesis. The second
	 task revisited the 2017 challenge by focusing on the discovery of word-like
	 units from unsegmented raw speech.
	 \item The Zero Resource Speech Challenge 2021~\cite{nguyen2020zero}, the latest
	 \edit{challenge}, expanded the scope to include language modeling tasks.
	 In addition to phoneme-level ABX, the challenge includes lexical,
	 semantic, and syntactic evaluation metrics computed via a language model of
	 pseudo-acoustic labels.
\end{itemize}
These challenges have facilitated the tracking of technical trends in
zero-resource speech technologies.
For example, research directions thereof 
have expanded to various speech processing components to cover the entire
spoken dialogue systems.
To keep up with this expansion, the challenge has continued to develop
appropriate evaluation metrics for zero-resource scenarios.
Following the success of representation learning, baseline and challenge
techniques have shifted from purely generative 
models~\cite{ondel2016variational,heck2017feature}, deep 
autoencoders~\cite{tjandra19_interspeech,chorowski_unsupervised_2019}, and incorporation of
neural-network-based TTS/VC techniques~\cite{niekerk20b_interspeech} to
self-supervised learning~\cite{maekaku21_interspeech}.
The latest challenge included the visual modality, continuing the
expansion to include more aspects of human interaction.

@@

Textless NLP is a new research direction that leverages the progress mentioned
above in self-supervised speech representation learning to model language
directly from audio, bypassing the need for text or 
labels~\cite{nguyen_generative_2022,polyak_speech_2021,kharitonov_textfree_2021,emotion_conversion}. 
Not only does this open the gate for language and dialect modeling without
orthographic rules, but it also offers the opportunity to model other
non-lexical information about how speech is delivered, e.g., speaker identity,
emotion, hesitation, interruptions. 
The generative spoken language model (nguyen_generative_2022)~\cite{nguyen_generative_2022} utilizes discrete
representations from wav2vec~2.0, HuBERT, and CPC algorithms as inputs to an
autoregressive language model trained by \edit{using the cross-entropy function} to maximize the
probability of predicting the next discrete speech token. A \edit{synthesis} module
follows the language model to produce speech waveforms given the generated
discrete speech units. The generated spoken continuations compete with
supervised generations and synthesis using a character language model in
subjective human evaluations. The model completes incomplete words
(pow[..] $\rightarrow$ POWER) and continues using words in the same general mood (dark $\rightarrow$
BLACKNESS)\footnote{https://speechbot.github.io/gslm/} and has been extended to
model and generate
dialogues~\cite{dialogue_nguyen_generative_2022}.\footnote{https://speechbot.github.io/dgslm/}
%\cite{}. 
Given its flexibility in modeling spoken content, the nguyen_generative_2022 has been further extended
to jointly model content and prosody~\cite{kharitonov_textfree_2021}. This prosodic-nguyen_generative_2022 model
introduced a multistream causal Transformer, where the input and output layers
use multiple heads to model three channels:  discrete speech units, duration,
and quantized pitch. The prosodic-nguyen_generative_2022 model jointly generates novel content
and prosody congruently in the expressive 
style of the prompt.\footnote{https://speechbot.github.io/pgslm/}
%\cite{}. 
Going one step further, \cite{emotion_conversion} used a speech emotion
conversion framework to modify the perceived emotion of a speech utterance
while preserving its lexical content and speaker identity. Other studies have
extended the idea of textless language processing or audio discrete representation to applications such as 
spoken question answering~\cite{spoken_qa_dual}, speech separation~\cite{shi2021discretization}, TTS~\cite{hayashi2020discretalk}, and speech-to-speech
translation~\cite{textless_translation}.
